Data assimilation for the wave equation with discontinuous coefficients

### 2024-03-26 tex/main.tex
Notes update

\pgfplotsset{
% initialize Set1-5:
cycle list/Set1-5,
% combine it with ’mark list*’:
cycle multiindex* list={
mark list*\nextlist
Set1-5\nextlist
},
}

@@

\section{Introduction}
\noindent In this article, we consider the discretization of a unique continuation problem for the wave equation with discontinuous coefficients. In the following, let $\Omega \subset \mathbb{R}^d$ be a bounded domain with smooth boundary $\partial \Omega$ such that there exists subdomains $\Omega_1, \Omega_2 \subseteq \Omega$ such that $\Omega = \overline{\Omega}_1 \cup \overline{\Omega}_2$ and $\Omega_1 \cap \Omega_2 = \emptyset$. We consider a scalar wave speed $c(x) = \mathbbm{1}_{\Omega_1} c_1(x) + \mathbbm{1}_{\Omega_2} c_2(x)$, $c_i \in C^\infty(\Omega_i)$, $i = 1,2$, which may be discontinuous across the interface $\Gamma = \Omega \setminus (\Omega_1 \cup \Omega_2)$, but constant in time. For a final time $T > 0$, we consider the space-time domain $Q = (0,T) \times \Omega$ with lateral boundary $\Sigma \coloneqq (0,T) \times \partial \Omega$. For a subset $\omega \subset \Omega$, we define the data domain $\omega_T \coloneqq (0,T) \times \omega$. In the following, let $\wop$ be the wave operator defined as 
\begin{equation*}
    \wop u := \partial_t^2 - \div(c(x)^2 \nabla u).
\end{equation*}
In particular, we write $\square_1$ if $c_i = 1$, $i = 1,2$, i.e. if we are considering the case of homogeneous media. 
Then, we consider the problem: find $u : Q \rightarrow \mathbb{R}$ such that 
\begin{equation}\label{eq:waveEquation}
        (\wop u, u) = (0,0) \quad \text{ in } Q \times \Sigma.  
\end{equation}
To ensure that the problem is well-posed, one usually prescribes initial data 
\begin{equation}\label{eq:initialData}
    (u,\dt u) \vert_{t = 0} = (u_0,u_1) \text{ in } \Omega. \tag{IVP} 
\end{equation}
Here we will consider a data assimilation problem instead, i.e. we assume that the initial data is unknown and that there are given measurements $u_{\omega} \in L^2(\omega_T)$ which are prescribed in $\omega_T$: 
\begin{equation}\label{eq:dataMatch}
    u = u_{\omega} \text{ in } \omega_T. \tag{DA}
\end{equation}

\noindent If $\partial \Omega$ is stricly convex and $c_i = 1$, $i = 1,2$, the problem \eqref{eq:waveEquation}-\eqref{eq:dataMatch} is Lipschitz stable if the set $\omega_T \subset Q$ fulfills the geometric control condition (GCC) \cite{BLR88,BLR92}. More precisely, we have the following theorem, cf. \cite[Thm. A.4 \& Rem. A.5]{BFMO21control}.

\begin{thm}\label{thm:Lipschitz}
    Assume that $c_i = 1$, $i = 1,2$, and that $\partial \Omega$ is stricly convex. If $\omega_T \subset Q$ fulfills the GCC, then there exists a constant $C > 0$ such that for any $\phi \in H^1(Q)$ we have the following estimates
    \begin{align*}
        \Vert \phi \Vert_{L^\infty(0,T;L^2(\Omega))} + \Vert \dt \phi \Vert_{L^2(0,T;H^{-1}(\Omega))} &\le C \left(  \Vert \phi \Vert_{L^2(\omega_T)} + \Vert \phi \Vert_{L^2(\Sigma)} + \Vert \wop \phi \Vert_{H^{-1}(Q)} \right), \\
        \Vert \phi \vert_{t = 0} \Vert_{L^2(\Omega)} + \Vert \dt \phi \vert_{t = 0} \Vert_{H^{-1}(\Omega)} &\le C \left(\Vert \phi \Vert_{L^2(\omega_T)} + \Vert \phi \Vert_{L^2(\Sigma)} + \Vert \wop \phi \Vert_{H^{-1}(Q)} \right). 
    \end{align*}
\end{thm}

@@

\noindent For $k \ge 1$, we define the $H^1$-conforming finite element space 
\begin{equation}
    V_h^k := \{ v \in H^1(\Omega) : v \vert_{K} \in \mathcal{P}^k(K) \ \forall K \in \mathcal{T}_h \},
\end{equation}
where $\mathcal{P}^k(K)$ denotes the space of polynomials of degree at most $k \in \mathbb{N}$ on $K \in \mathcal{T}_h$. Furthermore, we partition the time axis into $N$ subinterals $I_n = (t_n,t_{n+1})$, $n = 0, \dots, N-1$, where $0 = t_0 \le t_1 \le \dots \le t_N = T$. We assume that the intervals are of equal length and denote $\Delta t = \vert t_{n+1} -t_n \vert$. Consequently, we can define a corresponding partition of $Q$ and $\Sigma$ into time-slabs 
\begin{equation}
    \begin{aligned}
        Q^n := &I_n \times \Omega, \quad \Sigma^n := I_n \times \Sigma, \quad n = 0, \dots, N-1, \\
        Q = &\bigcup_{n = 0}^{N-1} Q^n, \quad \phantom{:} \Sigma = \bigcup_{n = 0}^{N-1} \Sigma^n.
    \end{aligned}
\end{equation}
In the following, we denote the space-time integrals on the time slabs as 
\begin{equation*}
    (u,v)_{Q^n} := \int_{I_n} \int_{\Omega} uv \ \dX \dT, \quad (u,v)_{\Sigma^n} := \int_{I_n} \int_{\Sigma} uv \ \dS \dT,
\end{equation*}
and define $\Vert v \Vert^2_{Q^n} := (v,v)_{Q^n}$ and $\Vert v \Vert^2_{\Sigma^n} := (v,v)_{\Sigma^n}$. 
Furthermore, we set 
\begin{equation*}
    \omega^n := I_n \times \omega, \quad (u,v)_{\omega^n} := \int_{I_n} \int_{\omega} uv \ \dX \dT, \quad \Vert v \Vert^2_{\omega_T} := \sum_{n = 0}^{N-1} (v,v)_{\omega^n}.
\end{equation*}
Finally, we define the time jump operator 
\begin{equation*}
    v^n_{\pm} (x) := \lim_{s \rightarrow 0^+} v(x,t_n \pm s), \quad \jump{v^n} := v^n_+ - v^n_-. 
\end{equation*}

@@

\noindent In the following, we denote
\begin{equation*}
    a(u,v)_{Q^n} := \int_{I_n} \int_{\Omega} c(x)^2 \nabla u \cdot \nabla v \ \dX \dT,
\end{equation*}
and introduce a bilinear form $A$ that represents the wave-equation \eqref{eq:waveEquation} in mixed formulation:  
\begin{equation}
    \begin{aligned}
        A[\Uh,\Yh] := \sum_{n = 0}^{N -1} \Big\{ &(\dt \ul_2, \yl_1)_{Q^n} + a(\ul_1,\yl_1)_{Q^n} + (\dt \ul_1 - \ul_2,\yl_2)_{Q^n} \\
        &- (c(x)^2 \nabla \ul_1 \cdot \mathbf{n}, \yl_1)_{\Sigma^n} \Big\},
    \end{aligned}
\end{equation} 
where $\Uh \in \ProdFullyDiscrSpace{k}{q}$ and $\Yh \in \ProdFullyDiscrSpace{k^\ast}{q^\ast}$ with $k, k^\ast, q \in \mathbb{N}$ and $q^\ast \in \mathbb{N}_0$. 
To approximate solutions of \eqref{eq:continuousProblem}, we search for stationary points of the Lagrangian 
\begin{align*}
    \cL_h (\Uh, \Zh) := &\frac{1}{2} \Vert \ul_1 - u_{\omega} \Vert^2_{\omega_T} + A[\Uh,\Zh] + \frac{1}{2} \gamma S_h(\Uh,\Uh) \\
    &- \frac{1}{2} \gamma^\ast S_h^\ast(\Zh,\Zh) + \frac{\gamma}{2} \Sud(\Uh,\Uh),
\end{align*}
where $S_h$, $S_h^\ast$, and $\Sud$ are stabilization terms yet to be defined. Note that the first and the second terms of $\cL_h$ incorporate the data and the PDE constraints, respectively. To define the stabilization terms $S_h$ and $S_h^\ast$, we introduce the following terms: 
\begin{align*}
    J(\Uh,\Wh) &:= \sum_{n = 0}^{N -1} \int_{I_n} \sum_{F \in \mathcal{F}_i} h (\jump{c(x)^2 \nabla \ul_1}, \jump{c(x)^2 \nabla \wl_1})_F \ \dT, \\
    G(\Uh,\Wh) &:= \sum_{n = 0}^{N -1} \int_{I_n} \sum_{K \in \mathcal{T}_h} h^2 (\dt \ul_2 - \div (c(x)^2 \nabla \ul_1),\dt \wl_2 -\div( c(x)^2 \nabla \wl_1))_K \ \dT, \\
    I_0(\Uh,\Wh) &:= \sum_{n = 0}^{N -1} (\ul_2 - \dt \ul_1, \wl_2-\dt \wl_1)_{Q_n}, \qquad 
    R(\Uh,\Wh) := \sum_{n = 0}^{N -1} h^{-1} (\ul_1,\wl_1)_{\Sigma^n}. 
\end{align*}
$J(\cdot,\cdot)$ is a continuous interior penalty term in space, $G(\cdot,\cdot)$ is a Galerkin least squares term enforcing the PDE locally on each element, $I_0(\cdot,\cdot)$ enforces that $\ul_2 = \dt \ul_1$, and $R(\cdot,\cdot)$ ensures boundary stability. 
With those terms, we define the primal stabilizer $S_h$ as 
\begin{equation}
    S_h(\Uh,\Wh) := J(\Uh,\Wh) + I_0(\Uh,\Wh) + G(\Uh,\Wh) + R(\Uh,\Wh), 
\end{equation}
and the dual stabilizer $S_h^\ast$ through 
\begin{equation}
    S_h^\ast(\Yh,\Zh) := \sum_{n = 0}^{N-1} \left\{ (\yl_1,\zl_1)_{Q^n} + a(\yl_1,\zl_1)_{Q^n} + (\yl_2,\zl_2)_{Q^n} + h^{-1} (\yl_1,\zl_1)_{\Sigma^n} \right\}.
\end{equation}
The stabilization term $\Sud$ imposes regularity on the discontinuities in time and is defined as 
\begin{equation}
    \Sud (\Uh,\Wh) := \underline{I}_1^{\uparrow \downarrow}(\Uh,\Wh) + \underline{I}_2^{\uparrow \downarrow}(\Uh,\Wh),
\end{equation}
where 
\begin{align*}
    \underline{I}_1^{\uparrow \downarrow}(\Uh,\Wh) &:= \sum_{n = 0}^{N-1} \left\{ \frac{1}{\Delta t} (\jump{\ul_1^n},\jump{\wl_1^n})_{\Omega} + \Delta t (c(x)^2 \jump{\nabla \ul_1^n},c(x)^2 \jump{\nabla \wl_1^n})_{\Omega}\right\}, \\
    \underline{I}_2^{\uparrow \downarrow}(\Uh,\Wh) &:= \sum_{n = 0}^{N-1} \frac{1}{\Delta t} (\jump{\ul_2^n},\jump{\wl_2^n})_{\Omega}.
\end{align*}

@@

\section{Analysis}
\noindent In this section, we analyze the approximation of the continuous problem \eqref{eq:continuousProblem} through stationary points of the Laplacian $\cL_h$. In particular, we show that the error between the first component $\ul_1$ of the discrete solution and the continuous solution $u$ is bounded by the best approximation error in suitable discrete norms defined below. Let $\vert \cdot \vert_{S_h}, \vert \cdot \vert_{\uparrow \downarrow}$, and $\Vert \cdot \Vert_{S_h^{\ast}}$ be the semi-norms, respectively norms, induced by the stabilizers $S_h$, $\Sud$, and $S_h^{\ast}$:
\begin{equation}
    \vert \Uh \vert^2_{S_h} := S_h(\Uh,\Uh), \quad \vert \Uh \vert^2_{\uparrow \downarrow} := \Sud(\Uh,\Uh), \quad \Vert \Zh \Vert^2_{S_h^{\ast}} := S_h^{\ast}(\Zh,\Zh).
\end{equation}
\noindent Then, we define the discrete norm 
\begin{equation}
    \tnorm{ (\Uh,\Zh) }^2 := \gamma \vert \Uh \vert^2_{S_h} + \vert \Uh \vert^2_{\uparrow \downarrow} + \Vert \ul_1 \Vert^2_{\omega_T} + \gamma^{\ast} \Vert \Zh \Vert^2_{S_h^\ast},
\end{equation}
and its strenghend version
\begin{equation}\label{eq:tnormwop}
    \begin{aligned}
        \tnorm{(\Uh, \Zh)}^2_{\wop} := \tnorm{(\Uh, \Zh)}^2 &+ \sum_{n = 0}^{N -1} \Big\{ \Vert \dt u_2 \Vert^2_{Q^n} + \Vert c^2 \nabla u_1 \Vert^2_{Q^n} + \Vert \dt u_1 \Vert^2_{Q^n} \\
        &+ \Vert u_2 \Vert^2_{Q^n} + \int_{I_n} \sum_{K \in \mathcal{T}_h} h^2 \Vert c^2 u_1 \Vert^2_{H^2(K)} \dT \Big\}. 
    \end{aligned}
\end{equation}
Note it suffices to show that $\tnorm{(\cdot,\cdot)}$ is a norm. We modify the steps from \cite{BP24} to show the following result.

\begin{lem}
    The expression $\tnorm{(\cdot,\cdot)}$ is a norm on $\ProdFullyDiscrSpace{k}{q} \times \ProdFullyDiscrSpace{k^\ast}{q^\ast}$.  
\end{lem}

@@

\noindent \textbf{Assume for now} that we can rewrite the estimates from Thm. \ref{thm:filippas} with $\Vert \wop u \Vert_{H^{-1}(Q)}$ instead of $\Vert \wop u \Vert_{L^2(Q)}$. As discussed in Rem. \ref{rem:homMedia}, we can then apply this continuum estimate to the residual $\tilde{e}_h := u - L_{\Delta t} \ul_1$, where the lifting operator $L_{\Delta t}$ ensures that $\tilde{e}_h \in H^1$. Following the argumentation from \cite[Thm. 4.4]{BP24}, it follows that 
\begin{equation}
    \Vert \ul_1 - L_{\Delta t} \ul_1 \Vert_{L^2(\omega_T)} \le C h^s \Vert u \Vert_{H^m(Q)}, 
\end{equation}
and $\Vert u - \ul_1 \Vert_{L^2(\omega_T)} \le \tnorm{(\mathbf{U}-\Uh,\Zh)}$ by definition of the triple-norm, it follows from Cor. \ref{cor:tnormConvRates} that $\Vert \tilde{e}_h \Vert_{L^2(\omega_T)} \le C h^s \Vert u \Vert_{H^m(Q)}$. Furthermore, \textbf{with similar arguments as in} \cite[Thm. 4.4]{BP24}, we obtain that 
\begin{equation}
    \Vert \wop \tilde{e}_h \Vert_{H^{-1}(Q)} \le C h^s \Vert u \Vert_{H^m(Q)}. 
\end{equation}
Thus, \eqref{eq:FilippasEstimateLog} yields 
\begin{equation}
    \Vert (u - \ul_1) \vert_{t = 0} \Vert_{L^2(\Omega)}  + \Vert \dt (u - \ul_1) \vert_{t = 0} \Vert_{H^{-1}(\Omega)} \le C C_{0,1} \log(1 + C_{0,1} h^{-s} \Vert u \Vert_{H^m(Q)}^{-1}),
\end{equation} 
where $C_{0,1} := \Vert (u_0,u_1) \Vert_{H^1 \times L^2}$. While this estimates holds under the general condition $T > 2 \sup_{x \in \Omega} d_c(x,\omega_T)$, we suspect that under the assumption of the geometric control condition, the estimate can be improved considerably. In particular, the numerical experiments considered in the following section suggest that a similar estimate to \eqref{eq:homMedia:estimate} holds in the heterogenous case.

@@

            \nextgroupplot[width=8cm,height=6.2cm,domain=1:4,xmode=linear,ymode=log, ylabel={}, xlabel={$L$}, title={$c_1 = 2.5$, $T = 0.5$, $\Vert \partial_t (u - u_h) \Vert_{L^2(0,T;L^2(\Omega_R))}$}, %cycle list name=, 
            legend pos=south west, %yticklabels={,,},
            ]
            %L2-L2-u_t (k = 2)
            \addplot+[discard if not={order}{2}, discard if not={contrast}{2.5},line width=1.1pt,color=teal,mark=*] table [x=L, y=L2-L2-error-u_t, col sep=comma] {../../wave_data_assimilation/data/jumpCoefs/simpleExact_T0.5_1D_jumpingCoefs_k3.csv};
            \addplot+[discard if not={order}{2}, discard if not={contrast}{2.5},line width=1.1pt,dashed,color=teal,mark=*] table [x=L, y=bestapprox-L2-L2-error-u_t, col sep=comma] {../../wave_data_assimilation/data/jumpCoefs/simpleExact_T0.5_1D_jumpingCoefs_k3.csv};

            %L2-L2-u_t (k = 3)
            \addplot+[discard if not={order}{3}, discard if not={contrast}{2.5},line width=1.1pt,color=red,mark=star] table [x=L, y=L2-L2-error-u_t, col sep=comma] {../../wave_data_assimilation/data/jumpCoefs/simpleExact_T0.5_1D_jumpingCoefs_k3.csv};
            \addplot+[discard if not={order}{3}, discard if not={contrast}{2.5},line width=1.1pt,dashed,color=red,mark=star] table [x=L, y=bestapprox-L2-L2-error-u_t, col sep=comma] {../../wave_data_assimilation/data/jumpCoefs/simpleExact_T0.5_1D_jumpingCoefs_k3.csv};

>>>>

%find colors at https://colorbrewer2.org/#type=qualitative&scheme=Set1&n=5
\pgfplotsset{
% initialize Set1-5:
cycle list/Set1-5,
% combine it with ’mark list*’:
cycle multiindex* list={
mark list*\nextlist
Set1-5\nextlist
},
}

@@

\section{Introduction}
\noindent In this article, we consider the discretization of a unique continuation problem for the wave equation with discontinuous coefficients. In the following, let $\Omega \subset \mathbb{R}^d$ be a bounded domain with smooth boundary $\partial \Omega$ such that there exists subdomains $\Omega_1, \Omega_2 \subseteq \Omega$ such that $\Omega = \overline{\Omega}_1 \cup \overline{\Omega}_2$ and $\Omega_1 \cap \Omega_2 = \emptyset$. We consider a scalar wave speed $c(x) = \mathbbm{1}_{\Omega_1} c_1(x) + \mathbbm{1}_{\Omega_2} c_2(x)$, $c_i \in C^\infty(\Omega_i)$, $i = 1,2$, which may be discontinuous across the interface $\Gamma = \Omega \setminus (\Omega_1 \cup \Omega_2)$, but constant in time. For a final time $T > 0$, we consider the space-time domain $Q = (0,T) \times \Omega$ with lateral boundary $\Sigma \coloneqq (0,T) \times \partial \Omega$. For any nonempty subset $\omega \subset \Omega$, we set $\omega_T \coloneqq (0,T) \times \omega$. In the following, let $\wop$ be the wave operator defined as 
\begin{equation*}
    \wop u := \partial_t^2 - \div(c(x)^2 \nabla u).
\end{equation*}
In particular, we write $\square_1$ if $c_i = 1$, $i = 1,2$, i.e. if we are considering the case of homogeneous media. 
Then, we consider the problem: find $u : Q \rightarrow \mathbb{R}$ such that 
\begin{equation}\label{eq:waveEquation}
        (\wop u, u) = (0,0) \quad \text{ in } Q \times \Sigma.  
\end{equation}
To ensure that the problem is well-posed, one usually prescribes initial data 
\begin{equation}\label{eq:initialData}
    (u,\dt u) \vert_{t = 0} = (u_0,u_1) \text{ in } \Omega. \tag{IVP} 
\end{equation}
Here we will consider a data assimilation problem instead, i.e. we assume that the initial data is unknown and that there are given measurements $u_{\omega} \in L^2(\omega_T)$ which are prescribed in $\omega_T$: 
\begin{equation}\label{eq:dataMatch}
    u = u_{\omega} \text{ in } \omega_T. \tag{DA}
\end{equation}

\noindent If $\partial \Omega$ is stricly convex and $c_i = 1$, $i = 1,2$, the problem \eqref{eq:waveEquation}-\eqref{eq:dataMatch} is Lipschitz stable if the set $\omega_T \subset Q$ fulfills the geometric control condition (GCC) \cite{BLR88,BLR92}. More precisely, we have the following theorem.

\begin{thm}\label{thm:Lipschitz}
    Assume that $c_i = 1$, $i = 1,2$, and that $\partial \Omega$ is stricly convex. If $\omega_T \subset Q$ fulfills the GCC, then there exists a constant $C > 0$ such that for any $\phi \in H^1(Q)$ we have the following estimates
    \begin{align*}
        \Vert \phi \Vert_{L^\infty(0,T;L^2(\Omega))} + \Vert \dt \phi \Vert_{L^2(0,T;H^{-1}(\Omega))} &\le C \left(  \Vert \phi \Vert_{L^2(\omega_T)} + \Vert \phi \Vert_{L^2(\Sigma)} + \Vert \square_1 \phi \Vert_{H^{-1}(Q)} \right), \\
        \Vert \phi \vert_{t = 0} \Vert_{L^2(\Omega)} + \Vert \dt \phi \vert_{t = 0} \Vert_{H^{-1}(\Omega)} &\le C \left(\Vert \phi \Vert_{L^2(\omega_T)} + \Vert \phi \Vert_{L^2(\Sigma)} + \Vert \square_1 \phi \Vert_{H^{-1}(Q)} \right). 
    \end{align*}
\end{thm}

\begin{proof}
    See \cite[Thm. A.4]{BFMO21control} and \cite[Rem. A.5]{BFMO21control}.
\end{proof}

@@

\noindent For $k \ge 1$, we define the $H^1$-conforming finite element space 
\begin{equation}
    V_h^k := \{ v \in H^1(\Omega) : v \vert_{K} \in \mathcal{P}^k(K) \ \forall K \in \mathcal{T}_h \},
\end{equation}
where $\mathcal{P}^k(K)$ denotes the space of polynomials of degree at most $k \in \mathbb{N}$ on $K \in \mathcal{T}_h$. Furthermore, we partition the time axis into $N$ subinterals $I_n = (t_n,t_{n+1})$, $n = 0, \dots, N-1$, where $0 = t_0 \le t_1 \le \dots \le t_N = T$. We assume that the intervals are of equal length and denote $\Delta t = \vert t_{n+1} -t_n \vert$. Then, we partition of $Q$ and $\Sigma$ into time-slabs 
\begin{equation}
    \begin{aligned}
        Q^n := &I_n \times \Omega, \quad \Sigma^n := I_n \times \Sigma, \quad n = 0, \dots, N-1, \\
        Q = &\bigcup_{n = 0}^{N-1} Q^n, \quad \phantom{:} \Sigma = \bigcup_{n = 0}^{N-1} \Sigma^n.
    \end{aligned}
\end{equation}
In the following, we denote the space-time integrals on the time slabs as 
\begin{equation*}
    (u,v)_{Q^n} := \int_{I_n} \int_{\Omega} uv \ \dX \dT, \quad (u,v)_{\Sigma^n} := \int_{I_n} \int_{\Sigma} uv \ \dS \dT,
\end{equation*}
and define $\Vert v \Vert^2_{Q^n} := (v,v)_{Q^n}$ and $\Vert v \Vert^2_{\Sigma^n} := (v,v)_{\Sigma^n}$. 
Furthermore, we set 
\begin{equation*}
    \omega^n := I_n \times \omega, \quad (u,v)_{\omega^n} := \int_{I_n} \int_{\omega} uv \ \dX \dT, \quad \Vert v \Vert^2_{\omega_T} := \sum_{n = 0}^{N-1} (v,v)_{\omega^n}.
\end{equation*}
Finally, we define the time jump operator 
\begin{equation*}
    v^n_{\pm} (x) := \lim_{s \rightarrow 0^+} v(x,t_n \pm s), \quad \jump{v^n} := v^n_+ - v^n_-. 
\end{equation*}

@@

\noindent In the following, we denote
\begin{equation*}
    a(u,v)_{Q^n} := \int_{I_n} \int_{\Omega} c(x)^2 \nabla u \cdot \nabla v \ \dX \dT,
\end{equation*}
and introduce a bilinear form $A$ that represents the wave-equation \eqref{eq:waveEquation} in mixed formulation:  
\begin{equation}
    \begin{aligned}
        A[\Uh,\Yh] := \sum_{n = 0}^{N -1} \Big\{ &(\dt \ul_2, \yl_1)_{Q^n} + a(\ul_1,\yl_1)_{Q^n} + (\dt \ul_1 - \ul_2,\yl_2)_{Q^n} \\
        &- (c(x)^2 \nabla \ul_1 \cdot \mathbf{n}, \yl_1)_{\Sigma^n} \Big\},
    \end{aligned}
\end{equation} 
where $\Uh \in \ProdFullyDiscrSpace{k}{q}$ and $\Yh \in \ProdFullyDiscrSpace{k^\ast}{q^\ast}$ with $k, k^\ast, q \in \mathbb{N}$ and $q^\ast \in \mathbb{N}_0$. 
To approximate solutions of \eqref{eq:waveEquation}-\eqref{eq:dataMatch}, we search for stationary points of the Lagrangian 
\begin{align*}
    \cL_h (\Uh, \Zh) := &\frac{1}{2} \Vert \ul_1 - u_{\omega} \Vert^2_{\omega_T} + A[\Uh,\Zh] + \frac{1}{2} \gamma S_h(\Uh,\Uh) \\
    &- \frac{1}{2} \gamma^\ast S_h^\ast(\Zh,\Zh) + \frac{\gamma}{2} \Sud(\Uh,\Uh),
\end{align*}
where $S_h$, $S_h^\ast$, and $\Sud$ are stabilization terms yet to be defined. Note that the first and the second terms of $\cL_h$ incorporate the data and the PDE constraints, respectively. To define the stabilization terms $S_h$ and $S_h^\ast$, we introduce the following terms: 
\begin{align*}
    J(\Uh,\Wh) &:= \sum_{n = 0}^{N -1} \int_{I_n} \sum_{F \in \mathcal{F}_i} h (\jump{c(x)^2 \nabla \ul_1}, \jump{c(x)^2 \nabla \wl_1})_F \ \dT, \\
    G(\Uh,\Wh) &:= \sum_{n = 0}^{N -1} \int_{I_n} \sum_{K \in \mathcal{T}_h} h^2 (\dt \ul_2 - \div (c(x)^2 \nabla \ul_1),\dt \wl_2 -\div( c(x)^2 \nabla \wl_1))_K \ \dT, \\
    I_0(\Uh,\Wh) &:= \sum_{n = 0}^{N -1} (\ul_2 - \dt \ul_1, \wl_2-\dt \wl_1)_{Q_n}, \qquad 
    R(\Uh,\Wh) := \sum_{n = 0}^{N -1} h^{-1} (\ul_1,\wl_1)_{\Sigma^n}. 
\end{align*}
$J(\cdot,\cdot)$ is a continuous interior penalty term in space, $G(\cdot,\cdot)$ is a Galerkin least squares term enforcing the PDE locally on each element, $I_0(\cdot,\cdot)$ enforces that $\ul_2 = \dt \ul_1$, and $R(\cdot,\cdot)$ ensures boundary stability. 
Then we define the primal stabilizer $S_h$ as 
\begin{equation}
    S_h(\Uh,\Wh) := J(\Uh,\Wh) + I_0(\Uh,\Wh) + G(\Uh,\Wh) + R(\Uh,\Wh), 
\end{equation}
and the dual stabilizer $S_h^\ast$ through 
\begin{equation}
    S_h^\ast(\Yh,\Zh) := \sum_{n = 0}^{N-1} \left\{ (\yl_1,\zl_1)_{Q^n} + a(\yl_1,\zl_1)_{Q^n} + (\yl_2,\zl_2)_{Q^n} + h^{-1} (\yl_1,\zl_1)_{\Sigma^n} \right\}.
\end{equation}
The remaining stabilization term $\Sud$ imposes regularity on the discontinuities in time and is defined as 
\begin{equation}
    \Sud (\Uh,\Wh) := \underline{I}_1^{\uparrow \downarrow}(\Uh,\Wh) + \underline{I}_2^{\uparrow \downarrow}(\Uh,\Wh),
\end{equation}
where 
\begin{align*}
    \underline{I}_1^{\uparrow \downarrow}(\Uh,\Wh) &:= \sum_{n = 0}^{N-1} \left\{ \frac{1}{\Delta t} (\jump{\ul_1^n},\jump{\wl_1^n})_{\Omega} + \Delta t (c(x)^2 \jump{\nabla \ul_1^n},c(x)^2 \jump{\nabla \wl_1^n})_{\Omega}\right\}, \\
    \underline{I}_2^{\uparrow \downarrow}(\Uh,\Wh) &:= \sum_{n = 0}^{N-1} \frac{1}{\Delta t} (\jump{\ul_2^n},\jump{\wl_2^n})_{\Omega}.
\end{align*}

@@

\section{Analysis}
\noindent In this section, we analyze the approximation of the continuous problem \eqref{eq:waveEquation}-\eqref{eq:dataMatch} through \eqref{eq:discreteProblem}. In particular, we show that the error between the first component $\ul_1$ of the discrete solution and the continuous solution $u$ is bounded by the best approximation error in suitable discrete norms defined below. Let $\vert \cdot \vert_{S_h}, \vert \cdot \vert_{\uparrow \downarrow}$, and $\Vert \cdot \Vert_{S_h^{\ast}}$ be the semi-norms, respectively norms, induced by the stabilizers $S_h$, $\Sud$, and $S_h^{\ast}$:
\begin{equation}
    \vert \Uh \vert^2_{S_h} := S_h(\Uh,\Uh), \quad \vert \Uh \vert^2_{\uparrow \downarrow} := \Sud(\Uh,\Uh), \quad \Vert \Zh \Vert^2_{S_h^{\ast}} := S_h^{\ast}(\Zh,\Zh).
\end{equation}
\noindent Then, we define the discrete norm 
\begin{equation}
    \tnorm{ (\Uh,\Zh) }^2 := \gamma \vert \Uh \vert^2_{S_h} + \vert \Uh \vert^2_{\uparrow \downarrow} + \Vert \ul_1 \Vert^2_{\omega_T} + \gamma^{\ast} \Vert \Zh \Vert^2_{S_h^\ast},
\end{equation}
and its strenghend version
\begin{equation}\label{eq:tnormwop}
    \begin{aligned}
        \tnorm{(\Uh, \Zh)}^2_{\wop} := \tnorm{(\Uh, \Zh)}^2 &+ \sum_{n = 0}^{N -1} \Big\{ \Vert \dt u_2 \Vert^2_{Q^n} + \Vert c^2 \nabla u_1 \Vert^2_{Q^n} + \Vert \dt u_1 \Vert^2_{Q^n} \\
        &+ \Vert u_2 \Vert^2_{Q^n} + \int_{I_n} \sum_{K \in \mathcal{T}_h} h^2 \Vert c^2 u_1 \Vert^2_{H^2(K)} \dT \Big\}. 
    \end{aligned}
\end{equation}
%Note it suffices to show that $\tnorm{(\cdot,\cdot)}$ is a norm. We modify the steps from \cite{BP24} to show the following result.

\begin{lem}
    The expressions $\tnorm{(\cdot,\cdot)}$ and $\tnorm{(\cdot,\cdot)}_{\wop}$ are norms on $\ProdFullyDiscrSpace{k}{q} \times \ProdFullyDiscrSpace{k^\ast}{q^\ast}$.  
\end{lem}

@@

\noindent \textbf{Assume for now} that we can rewrite the estimates from Thm. \ref{thm:filippas} with $\Vert \wop u \Vert_{H^{-1}(Q)}$ instead of $\Vert \wop u \Vert_{L^2(Q)}$. As discussed in Rem. \ref{rem:homMedia}, we can then apply this continuum estimate to the residual $\tilde{e}_h := u - L_{\Delta t} \ul_1$, where the lifting operator $L_{\Delta t}$ ensures that $\tilde{e}_h \in H^1$. Following the argumentation from \cite[Thm. 4.4]{BP24}, it follows that 
\begin{equation}
    \Vert \ul_1 - L_{\Delta t} \ul_1 \Vert_{L^2(\omega_T)} \le C h^s \Vert u \Vert_{H^m(Q)}, 
\end{equation}
and $\Vert u - \ul_1 \Vert_{L^2(\omega_T)} \le \tnorm{(\mathbf{U}-\Uh,\Zh)}$ by definition of the triple-norm, it follows from Cor. \ref{cor:tnormConvRates} that $\Vert \tilde{e}_h \Vert_{L^2(\omega_T)} \le C h^s \Vert u \Vert_{H^m(Q)}$. Furthermore, \textbf{with similar arguments as in} \cite[Thm. 4.4]{BP24}, we obtain that 
\begin{equation}
    \Vert \wop \tilde{e}_h \Vert_{H^{-1}(Q)} \le C h^s \Vert u \Vert_{H^m(Q)}. 
\end{equation}
Thus, \eqref{eq:FilippasEstimateLog} yields 
\begin{equation}
    \Vert (u - \ul_1) \vert_{t = 0} \Vert_{L^2(\Omega)}  + \Vert \dt (u - \ul_1) \vert_{t = 0} \Vert_{H^{-1}(\Omega)} \le C C_{0,1} \log(1 + C_{0,1} h^{-s} \Vert u \Vert_{H^m(Q)}^{-1})^{-1},
\end{equation} 
where $C_{0,1} := \Vert (u_0,u_1) \Vert_{H^1 \times L^2}$. While this estimates holds under the general condition $T > 2 \sup_{x \in \Omega} d_c(x,\omega_T)$, we suspect that under the assumption of the geometric control condition, the estimate can be improved considerably. In particular, the numerical experiments considered in the following section suggest that a similar estimate to \eqref{eq:homMedia:estimate} holds in the heterogenous case.

@@

            \nextgroupplot[width=8cm,height=6.2cm,domain=1:4,xmode=linear,ymode=log, ylabel={}, xlabel={$L$}, title={$c_1 = 2.5$, $T = 0.5$, $\Vert \partial_t (u - u_h) \Vert_{L^2(0,T;L^2(\Omega_R))}$}, %cycle list name=, 
            legend pos=south west, %yticklabels={,,},
            ]
            %L2-L2-u_t (k = 2)
            \addplot+[discard if not={order}{2}, discard if not={contrast}{2.5},line width=1.1pt,color=orange,mark=square*] table [x=L, y=L2-L2-error-u_t, col sep=comma] {../../wave_data_assimilation/data/jumpCoefs/simpleExact_T0.5_1D_jumpingCoefs_k3.csv};
            \addplot+[discard if not={order}{2}, discard if not={contrast}{2.5},line width=1.1pt,dashed,color=orange,mark=square*] table [x=L, y=bestapprox-L2-L2-error-u_t, col sep=comma] {../../wave_data_assimilation/data/jumpCoefs/simpleExact_T0.5_1D_jumpingCoefs_k3.csv};

            %L2-L2-u_t (k = 3)
            \addplot+[discard if not={order}{3}, discard if not={contrast}{2.5},line width=1.1pt,color=cyan!60!black,mark=diamond*] table [x=L, y=L2-L2-error-u_t, col sep=comma] {../../wave_data_assimilation/data/jumpCoefs/simpleExact_T0.5_1D_jumpingCoefs_k3.csv};
            \addplot+[discard if not={order}{3}, discard if not={contrast}{2.5},line width=1.1pt,dashed,color=cyan!60!black,mark=diamond*] table [x=L, y=bestapprox-L2-L2-error-u_t, col sep=comma] {../../wave_data_assimilation/data/jumpCoefs/simpleExact_T0.5_1D_jumpingCoefs_k3.csv};
